---
title: "Environmental behaviour & practices"
description: |
  some
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
repository_url: https://github.com/LF2L/everest-bio
#bibliography: 03-comportement.bib
bibliography: library.bib
#nocite: '@*'
  
author:
  - first_name: " Master 3"
    last_name: ""
    url: 
    affiliation: CREIDD?, ERPI?
    affiliation_url: 
    
  - first_name: "Fabio"
    last_name: "Cruz"
    url: https://erpi.univ-lorraine.fr/people/Fabio-Cruz/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0001-8529-5327    
    
  - first_name: "Nadège"
    last_name: "Troussier"
    url: https://recherche.utt.fr/icd-directory/nadege-troussier
    affiliation: CREIDD
    affiliation_url: https://recherche.utt.fr/icd-directory/nadege-troussier
    orcid_id: 0000-0001-6707-0300

  - first_name: "Hakim"
    last_name: "Boudaoud"
    url: https://erpi.univ-lorraine.fr/people/Hakim-Boudaoud/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0003-4452-5952
    
  - first_name: "Mauricio"
    last_name: "Camargo"
    url: https://erpi.univ-lorraine.fr/people/Mauricio-Camargo/
    affiliation: ERPI
    affiliation_url: https://erpi.univ-lorraine.fr/
    orcid_id: 0000-0003-3867-2438 
always_allow_html: true        
---
```{r, include=FALSE}
require(DiagrammeR)
require(tidyverse)
require(kableExtra)
require(widgetframe)
```

```{r, include=FALSE}

#library(rticles)
#library(rmarkdown)
#render("03-Sustainable-behaviour.Rmd", output_format = "word_document", output_dir = "03-Sustainable-behaviour")


#render("01-Sustainable-indicators.Rmd", output_format = "pdf_document", output_dir = "01-Sustainable-Indicators")
#render("01-Sustainable-indicators.Rmd", output_format = "elsevier_article", output_dir = "01-Sustainable-Indicators")

```

# Abstract

Master 3: Mise en place de la collecte des bouchons : pourrait être analyse sur le comportement des acteurs => quel changement dans ces profils de gens ? => autre master (identifier ce que ça change chez les gens qui participent ou pas) avec Tatiana et Laure !!! doctorant avec master côté Nancy (avec Hakim et Mauricio) 


[**Link to modify the Graph**](https://lucid.app/lucidchart/invitations/accept/f6847ab0-3fd4-4749-a581-669058760d2d)


:::l-page
<div style="width: 100%; height: 480px; margin: 10px; position: relative; text-align: center;"><iframe allowfullscreen frameborder="0" style="width:95%; height:480px" src="https://lucid.app/documents/embeddedchart/945c1be7-a621-4c3e-9267-f77f2fb2855a" id="rc17nDan15JX"></iframe></div>
:::

# Rationale

Some ideas:

## On Circular Economy (CE)

- Circular economy decisions are very complex but often the perceptions, attitudes, and personal values of decision makers directly influence organizations’ strategic actions towards a circular economy. Both active participation of organizations and the role of decision makers are pivotal to implement a circular economy for plastics [@Khan2020].

- Identifying and assessing specific barriers associated with the workplace context could be crucial for achieving better employee engagement and a lower ecological footprint for organizations [@Yuriev2020]

### Barriers of CE implementation

- @Khan2020 identified  barriers to deploy CE Plastics in order of importance are as follows. The foremost barrier is that organizations (Entreprises) do not have easy access to recycling facilities. 
Second, organizations cannot transport plastic waste to recycling facilities. 
Third, organizations do not have enough space to store plastic waste. 
Fourth, organizations do not have enough time to sort plastic waste. 
Fifth, organizations do not have sufficient funds to pursue plastic recycling. 
Sixth, organizations do not have skilled personnel to sort plastic waste. 
Seventh, organizations do not have an environmental expert at the management level. 
Lastly, organizations do not know about potential buyers of plastic waste. 
These findings imply that prominent barriers are physical factors rather than knowledge. Overall, these findings contribute to the literature on the barriers of a circular economy

## On Theory of planned Behaviour


- @Ajzen1991 developed **Theory of Planned Behaviour (TPB)**  which is one of the most influencing model for designing behavioral interventions [@Yuriev2020]. Therefore, researchers have been increasingly employing the TPB model across various disciplines including environmental science and sustainability management
TPB postulates considers three main aspects 

1.  Attitude towards behavioral intention, 
2. Subjective Norms (SN), and 
3. Perceived Behavioral Control (PCB),  

together they shape behavioral intention as well as actual behavior.  To date, TPB has received around 90,000 citations [@Yuriev2020]

```{r, include=TRUE, echo=FALSE}

mermaid("
graph LR
  A[Attitude toward the act <br/> -ATT-]-->D[Behavioural Intention <br/>]
  B[Subjective Norm -SN-<br/>]-->D[Behavioural Intention]
  C[Perceived Behavioural Control -PBC-<br/>]-->D[Behavioural Intention]
  D[Behavioural Intention <br/>]-->E[Behaviour]
  C[Perceived Behavioural Control -PCB-]-.->E[Behaviour]
 
  classDef className font-size:18px, height:60px ;
  class A,B,C,D className;
")

#
#  
```




-	Theory of planned behavior (TPB) is based on the assumption that some conscious reasoning is involved in the formation of intentions to perform a behavior.
  - Behavior partly on the control of the individual
  -	Behavior is predicted by attitudinal and normative factors and perceived behavioral control (PBC). 
-	PBC reflects the extent to which the individual feels able to perform the behavior. 
-	Past behavior tends to predict intentions and future behavior
-	The factors that influenced the past behavior continue to influence the intentions and future behavior, but past behavior does not cause future behavior.


```{r, include=FALSE}
TPB <- 
tribble(
  ~Variable, ~Definition, ~Definition.recycling,
  "Attitude", 
    "Attitude toward the behavior and refers to the degree to which a person has a favorable or unfavorable evaluation or appraisal of the behavior in question",
      "The extent to which plastic recycling is valued by people, whether positive or negative",
  
  "Subjective Norms",
   "The perceived social pressure to perform or not to perform the behavior",
      "The extent to which people are influenced by perceived norms of plastic recycling in their  surroundings",
      
  "Perceived Behavioral Control (PCB)",   
    "The perceived ease or difficulty of performing the behavior and it is assumed to reflect past experience as well as anticipated impediments and obstacles", 
        "the perceived power (and knowledge) of people to implement plastic recycling",
  
  "Behavioral intention",   
    "They are indications of how hard people are willing to try, of how much of an effort they are planning to exert, in order to perform the behavior",
    "A perceived likelihood or subjective probability of decision makers that they will implement plastic recycling in their organizations",
  
  
  "Behaviour", 
    " ",
      "The extent to which organizations have implemented best practices of plastic recycling"
)
```


```{r, include=TRUE, layout="l-page", echo=FALSE}

kbl(TPB) %>%
  kable_paper(full_width = F)
```


## On Distributed Recycling via Additive Manufacturing

- The centralized schemes is the current paradigm for recycling concerns large majority of research on recycling behaviour comes within the scope of municipal waste management schemes. 

- Centralized recycling schemes have been criticized for inducing much waste transportation, which limits the ecological and economic benefits of recycling, especially in low density areas. 
Therefore, in order to mitigate CO2 emissions and financial costs of transporting, the concept of distributed recycling emerged. Distributed recycling schemes are characterized by decentralized plastic waste collection, sorting and transformation. 
It includes in-house recycling, or recycling in local small-scale recycling networks, be it in a community, a company, or a public organisation. If understandings of distributed recycling regarding the decentralisation level differs, scholars agree on its potential for reducing gas emissions from waste collection  and creating groundbreaking alternative value chains to the established raw-resources based manufacturing . 

For Kreiger (2014), distributed recycling schemes could bring benefits in any economic context, independently of development levels. In the present study, the technology enabling recycling in a distributed recycling scheme is 3D printing, which we call distributed recycling via additive manufacturing (DRAM). A large variety of goods can be produced recycling plastic with 3D technology. Therefore, we believe DRAM has the potential for enhancing the circular economy approach and “close the loop”. 





# On pro-environmental behaviours

- @Yuriev2020 identified nine systematic literature reviews on pro-environmental behaviors  (Li et al., 2019; Lo et al., 2012; Koop et al., 2019; Staddon et al., 2016; Varotto and Spagnolli, 2017; Inoue and Alfaro-Barrantes, 2015; Francoeur et al., 2019; Norton et al., 2015; Yuriev et al., 2018). 
These are the only existing systematic literature reviews from the last 10 years that covered various individual environmental actions. Having


- Chan, L., Bishop, B., 2013. A moral basis for recycling: Extending the theory of planned behaviour. J. Environ. Psychol. 36, 96–102. https://doi.org/10.1016/j.jenvp.2013.07.010







