---
title: "Everest Bio"
subtitle: "Ecosystem services for Distributed recycling via Additive Manufacturing"  
author: 
  - "Fabio CRUZ"
  - ""
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)
# Link for the Figures
URL = c('https://raw.githubusercontent.com/fabbiocrux/Figures/main/')

#library(xaringanBuilder)
#build_pdf("ES-presentation.Rmd")



```

```{r, include=FALSE, echo=FALSE}
# CSS for the CSS
htmltools::includeCSS("https://raw.githubusercontent.com/fabbiocrux/presentaciones/gh-pages/Assets-global/css/Fabio-fonts.css")
htmltools::includeCSS("https://raw.githubusercontent.com/fabbiocrux/presentaciones/gh-pages/Assets-global/css/Fabio.css")
```


class: title-slide, right, middle
background-position: left
background-size: contain

.pull-right[

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$subtitle`

### `r rmarkdown::metadata$author`

### `r rmarkdown::metadata$date`
]


---

# Agenda

1. Ecosystem Services 
2. ES and Distributed Recycling
3. Tool 


--- 

# Definition

---

## Types of ES

--
## Barriers and complexities of ES
- Adoption of ES in decision-making processes, for example data availability and accuracy, and lack of resources (time, competences, and money) to produce the assessments [@Longato2021]
- “the multiple and overlapping scales of ES” in a plan at the urban scale, considering also that ES knowledge is not always easily translatable into specific spatial units
